Regional Inequalities and economic growth; cross section and times series analysis

### 2023-11-02 MSB104_GR_1_Final_Assignment_research_article.qmd
Add files via upload

Using time series data and cross sectional observations we investigate the GDP per capita and GINI trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020. The aims is to acquire, process, and analyze sub-national GDP and population data for a subset of European countries.

The primary objective is to gain knowledge of econometric methods used for science in business through our given topic economic growth and inequality measures. Both GDP per capita and the Gini coefficient are vital indicators in the field of economics and are frequently used in a wide array of studies to analyse and interpret the economic conditions and social disparities of different regions or countries. Their analysis provides crucial insights for policymakers, economists, and researchers in shaping and evaluating the effectiveness of socio-economic policies.

The research question ***"What are the regional economic growth and inequality measurements from 2000 to 2020, in selected European countries and how does transport infrastructure affect the results?"*** is key in guiding our study. Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also critically examines the role of transportation infrastructure play in these dyniamics. @lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality transportation costs have the potential to reduce regional inequalities.

The sructure of our study is methodically designed to facilitate this exploration. In part 1A and B we calculate GDP per capita and explore regional inequity using various descriptive statistics and visualizations such as panel data (***Assignment_1***).

In part 2A and B we use cross sectional method to examining the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).

# Literature Review: Regional Economic Growth, Inequality, and Transport Infrastructure (2000-2020)

Why some places grow and prosper is a fundamental question in social science research. This question motivated Adam Smith\'s work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A. (1863). An inquiry into the nature and causes of the wealth of nations. Edinburgh: A. & C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states. Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Lessmann and Seidel\'s research offers insights into these dynamics, highlighting the persistence of income disparities across regions despite overarching economic advancement [@lessmann2017]. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.\'s findings [@gennaioli2014]. However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia showcasing different growth trajectories, particularly after their EU accession.

@@

The 2008 financial crisis significantly disrupted regional growth and exacerbated income inequalities across Europe. Nguyen et al. highlight that financial turmoil typically leads to increased income disparity, a trend observed globally [@nguyen2022].

Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts. Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential acceleration in growth for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014]. However, the financial crisis and subsequent labor market shifts may have influenced these trends.

@@

@lessmann2017 regional inequality study address the importance of studying regional inequality. In particular, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017]. The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.

@@

In Lessmann and Seidel\'s exploration of regional inequality, the intricacies of this phenomenon unfold, revealing a multifaceted landscape. The study delineates how economic development carves an N-shaped trajectory in regional disparities, illustrating an initial surge in inequality, followed by a decrease, and an eventual uptick again, paralleling the region\'s economic progress. Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].

Urbanization emerges as a crucial determinant, where bustling, populous urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017]. The regulatory landscape too plays a formidable role; robust market regulations, particularly within the realms of capital and labour, are highlighted as vital catalysts fostering regional convergence and mitigating the chasms of inequality [@lessmann2017]. Collectively, these determinants underscore the complexity of regional inequality, prompting a call for comprehensive and nuanced strategies to navigate and address its multifarious dimensions.

*Transport infrastructure*

Transport infrastructure plays a pivotal role in shaping regional economic landscapes. Improved connectivity fosters economic activities, reduces travel time, and enhances accessibility to markets and resources. Studies indicates that regions with robust transport infrastructure tend to attract more investments, generate employment, and ultimately contribute to regional prosperity.

@@

Furthermore, the paper provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017]. The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.

@@

@iammarino2019 discusses regional inequalities in European countries. It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium. The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019]. Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of under performance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019]. On the other hand, France has been a subject of discussion and policy making for many years, regarding its pockets of extreme wealth. (finner ikke kilde)

Lastly, It is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure emerging as a significant determinant. Earlier studies provides us with a foundational understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities. The integration of transport infrastructure into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.

@@

Trough Eurostat, we download the datasets nama_10r_3gdp and demo_r_pjanggr3 as csv files and filted the data by our preferences before downloading it. Furthermore, we filtered the dataset by choosing the years 2000 to 2020. Then, we selected the NUTS 3 region for Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified the data to be in million Euro.

### GDP

The GDP dataset provides insights into GDP at regional level using the NUTS classification system. It furnishes GDP values in both current prices and adjusted for inflation, with figures given in purchasing power standards (PPS) to account for price level differences between countries. The data is mostly structured by year and region @eurostat2023.

The GDP at market prices represents the final result of production activities of resident producer units within a region or nation. It is calculated as the sum of the gross value added across various institutional sectors or industries. Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @eurostat2023c. This also balances out in the total economy production account. In terms of methodology, while national accounts compile GDP from the expenditure side, regional accounts don't adopt this approach due to the complexities of accurately mapping inter-regional flows of goods and services.

The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @eurostat2023c.

### Population

Eurostat's records annual population data with NUTS classification. Our dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender. Displayed in a year-by-region format, with yearly interval updates @eurostat2023a.

Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project. The project covers 37 countries and is the central repository for demographic and migration-related data @eurostat2023a. As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @eurostat2023a. Additionally, data on marriages, divorces, and migration flows are recorded .

For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @eurostat2021.

### NUTS classification

@@

### Descriptive statistics

@@

# Part 1A: Sub-national GDP and GDP per Capita

## GDP per Capita Calculation

@@

### Portugal

@@

By looking at figure for Portugal, we can see that the GDP per capita in Portugal's regions appears to be fairly consistent. There is however some regional variability. We can see that the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas. Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).

@@

### France

@@

### Hungary

@@

### Slovakia

@@

### Denmark

@@

# Part 1B: Regional Inequity

## Gini Coefficient Calculation

@@

we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.

@@

# Part 2A: Growth and Inequity

@@

@@

**###Vis formel##**

Y is the dependent variable (hva som perdikeres).

X is the independent variable (input).

β0​ is the intercept.

β1​ is the slope of the line.

ε represents the residuals or error in the prediction.

-   **Intercept (**β0​): represents the value of Y when X is 0.

@@

"In linear regression, the objective is straightforward: determine the line that best represents the observed data. Furthermore, this 'goodness of fit' line is characterized by coefficients β0​ and β1, which represent the intercept and slope, respectively [@wooldridge2020]. The optimal line is one that minimizes the sum of the squared differences between the observed values and the predictions made by the model [@wooldridge2020]. This approach, known as the Ordinary Least Squares (OLS), has a solid mathematical foundation [@wooldridge2020]. By taking the sum of these squared differences and setting their partial derivatives with respect to β0​ and β1 to zero, we can derive the formula that give us these coefficients. And once these coefficients are in hand, they provide a direct equation for the line that best fits our data set [@wooldridge2020]. One of the strengths of OLS is its closed-form solution, meaning we can directly compute the coefficients from the data without iterative procedures [@wooldridge2020]. Moreover, under the foundamental assumptions of linear regression, OLS stands out as the Best Linear Unbiased Estimator (BLUE), emphasizing its reliability and efficiency in estimating the true underlying relationship [@wooldridge2020]."

### Individual country cases

#### Portugal

#### Denmark

#### France

#### Hungary

#### Slovakia

### **Outline discussion**

**##Eksepel....When looking at the data it is obvious that we have too many regions with zero inequality. This seems strange. The reason for this is ..... In the subsequent analysis it therefore would make scene to exclude these observations to not bias our finding.**

**Looking at specific countries we observe allso further abnormalities like in .... where ... As thees outliers are not do to a bias in our data we shold keep them in our data. However we need to test later the sensitivity of our findings regarding the presences of these borderline observations. ...####**

# Part 2B: Exploring Other Determinants of Inequity

@@

@@

>>>>

Using time series data and cross-sectional observations we investigate Economic growth and Inequality trends in Portugal, France, Hungary, Slovenia and Denmark in the time period 2000 to 2020. The aims are to acquire, process, and analyse sub-national GDP, GINI and population data for a subset of European countries.

The primary objective is to gain knowledge of econometric methods used for science in business through our given topic. Subsequently, econometric terms and definitions are explained in greater detail than it normally would in similar papers. Using GDP per capita and the Gini coefficient. Moreover, both are vital indicators in the field of economics and are frequently used, in a wide array of studies. Furthermore, to analyse and interpret the economic conditions and social disparities of different regions or countries. Their analysis provides crucial insights for policymakers, economists, and researchers in shaping and evaluating the effectiveness of socio-economic policies.

The research question ***"What are the regional economic growth and inequality measurements from 2000 to 2020, in selected European countries and how does other*** ***determinants affect the results?"*** is key in guiding our study. Setting a solid foundation for a comprehensive study that not only measure our given parameters, but also a critically examination of the role of determinants such as transportation infrastructure and education. @lessmann2017 states that transportation costs play an important role in agglomeration and income, and reduced transportation costs as a determinant of regional inequality.

The sructure of our study is methodically designed to facilitate this exploration. In part 1A and B we calculate GDP per capita and explore regional inequality using various descriptive statistics and visualizations such as panel data (***Assignment_1***).

In part 2A and B we use cross sectional method to examine the relationship between regional development and inequality, employing cross sectional estimation techniques for the year 2010 (***Assignment_2***).

Lastly, potential limitations may arise due to the use of ChatGPT and ChatGPT PDF, which have been utilized as tools to gain understanding and extract knowledge from academic papers, as well as to explain econometric terms and definitions. The text generated by these tools has been cross-checked to the best of our knowledge with our textbook and Andre Seidel, lecturer and author of 'Regional Inequality, Convergence, and Its Determinants -- A View from Outer Space' (2017). Despite these precautions, there may have been instances where incorrect information was provided, and some errors could have been overlooked. The text has been meticulously rewritten to fit our style of writing, ensuring consistency and accuracy throughout our study.

# Literature Review: Regional Economic Growth, Inequality and determines (2000-2020)

Why some places grow and prosper is a fundamental question in social science research. This question motivated Adam Smith's work, *An Inquiry into the Nature and Causes of the Wealth of Nations,* (Smith [1863](https://link-springer-com.galanga.hvl.no/article/10.1007/s11187-014-9574-4#ref-CR32 "Smith, A. (1863). An inquiry into the nature and causes of the wealth of nations. Edinburgh: A. & C Black.")) and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. In recent decades, the European Union has implemented policies aimed at reducing regional disparities and promoting economic growth across its member states. Furthermore, despite these efforts, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe exhibits a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Lessmann and Seidel's research offers insights into these dynamics, highlighting the persistence of income disparities, across regions despite overarching economic advancement [@lessmann2017]. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by Gennaioli et al.'s findings [@gennaioli2014]. However, this is not uniformly experienced across all European countries, with Eastern European nations such as Hungary and Slovakia we may expect different growth trajectories, particularly after their EU accession.

@@

The 2008 financial crisis significantly disrupted regional growth and exacerbated income inequalities across Europe. Nguyen et al. highlight that financial turmoil typically leads to increased income disparity; a trend observed globally [@nguyen2022].

Focusing on Europe, we expect varied regional economic performances, especially when comparing Eastern European countries such as Hungary and Slovakia to their Western counterparts. Given the role of capital market regulations in promoting regional convergence, outlined by Gennaioli et al., we anticipate a potential growth acceleration. In particular for Eastern European countries following their EU accession in 2003 and 2004 [@gennaioli2014]. However, the financial crisis and subsequent labour market shifts may also have influenced these trends.

@@

@lessmann2017 regional inequality study address the importance of studying regional inequality. Its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017]. The paper also discusses the relationship between regional inequality and personal income inequality and conflict risk.

@@

Geographic characteristics stand out as pivotal, with the natural and topographical elements of a region significantly swaying income distributions [@lessmann2017].

Urbanization emerges as a crucial determinant, where bustling, populous around urban centres usually enjoy elevated income levels, contrasting starkly with the often-stark disparities witnessed in more rural and secluded areas [@lessmann2017].

*Transport infrastructure and Education*

Transport infrastructure plays a central role in shaping regional economic landscapes. Improved connectivity fosters economic activities, reduces travel time, and enhances accessibility to markets and resources. Studies indicates that regions with robust transport infrastructure tend to attract more investments, generate employment, and ultimately contribute to regional prosperity.

[@lessmann2017] discusses the determinant of Education in the context of human capital. The study mentions that human capital is the most important determinant of differences in regional development within countries. The study also notes that human capital creates positive externalities, which can spill over regional boundaries. Moreover, the quality of human capital within countries, as measured by the secondary-school enrolment rate, can decrease regional inequality by facilitating regional spillovers and convergence. Lastly, promoting internal migration.

@@

Furthermore, [@lessmann2017] provides us with insights into the inverted U-shaped relationship between regional inequality and the level of economic development in different country groups [@lessmann2017]. The inverted U-shape also referred to as Kuznets Curve suggests that as an economy develops, regional inequality initially rises, reaches a peak, and then starts to decline.

@@

@iammarino2019 discusses regional inequalities in European countries. It presents evidence that inter-regional inequality has been increasing in the European Union since the turn of the millennium. The authors argue that this is due to the existence of several groups of regional economies in Europe that are structurally very different from one another [@iammarino2019]. Additionally, the authors argue that there are some countries in the EU that are more evenly developed than others, and that a map of underperformance or over-performance means less in a high-income but evenly developed country, such as in Denmark in our case [@iammarino2019]. On the other hand, France has been a subject of discussion and policy making for many years, regarding its well-known pockets of extreme wealth in regions such as Cap d'Antibes.

Overall, it is evident that regional economic growth and inequality in Europe from 2000 to 2020 have been influenced by a constellation of factors, with transport infrastructure and education emerging as significant determinants. Earlier studies provide us with a fundamental understanding of these dynamics, while also highlighting the nuanced impacts of economic crises on regional disparities. The integration of transport infrastructure and education into this discourse adds an additional layer of complexity, underscoring its potential as a catalyst for regional development and equality.

@@

Trough Eurostat, we download the datasets nama_10r_3gdp , demo_r_pjanggr3, demo_r_mlifexp, tran_r_net and edat_lfse_04 as csv files and filted the data by our preferences before downloading it. Furthermore, we filtered the dataset by choosing the years 2000 to 2020. Then, we selected the NUTS requiered levels region for Portugal, France, Hungary, Slovakia and Denmark. Finally, we specified the data to be in million Euro.

## GDP

The GDP dataset provides insights into GDP at regional level using the NUTS classification system. It furnishes GDP values in both current prices and adjusted for inflation, with figures given in purchasing power standards (PPS) to account for price level differences between countries. The data is mostly structured by year and region @database.

The GDP at market prices represents the final result of production activities of resident producer units within a region or nation. It is calculated as the sum of the gross value added across various institutional sectors or industries. Furthermore, augmented by taxes and reduced by subsidies on products (which are not allocated to specific sectors or industries) @database. This also balances out in the total economy production account. In terms of methodology, while national accounts compile GDP from the expenditure side, regional accounts don't adopt this approach due to the complexities of accurately mapping inter-regional flows of goods and services.

The different measures for the regional GDP are absolute figures in € and Purchasing Power Standards (PPS), figures per inhabitant and relative data compared to the EU Member States average @database.

## Population

Eurostat's records annual population data with NUTS classification. Our dataset includes information on births, deaths, net migration, and may also include demographic information on age and gender. Displayed in a year-by-region format, with yearly interval updates @database.

Eurostat's primary source for yearly demographic data at the regional level stems from the Unified Demography (Unidemo) project. The project covers 37 countries and is the central repository for demographic and migration-related data @database. As well as, specific metrics gathered under UNIDEMO encompass population counts at the close of the calendar year and events such as births and deaths occurring within that year @database. Additionally, data on marriages, divorces, and migration flows are recorded .

For the purpose of this research, the demographic data references the NUTS 2016 classification, which provides a detailed breakdown of the European Union's territory @database.

## Demographic (Nuts 2)

Statistics on population change (demo_r_mlifexp) and on population structure was collected from Euro Stat [@database]. Used to monitor demographic behavior within political, economic, social and cultural contexts at NUTS 2 level. Same policy and demographic classifications as previous population data set applies.

## Transportation Infrastructure (Nuts 2)

This data-set (tran_r_net) presents recent data on the inland transport network of the European Union (EU), EFTA and candidate countries presenting motorways, railways and inland waterways [@database]. Furthermore used as a varable of of interest (MLR) in our section growth and inequality. Moreover, to excamine if evolution of the transport network is closely linked to the general development of the economy.

The collected data are compiled by the competent national authorities. Reported annually be the National Statistical Office, the Civil Aviation Authorities (air transport), the Ministry of Transport (inland waterways, railway and road networks, victims), and the National Maritime Administration (maritime transport) [@database].

## Education (Nuts 2)

This data set (edat_lfse_04) contains population by educational attainment level, sex and NUTS 2 regions (%). Including data on the highest level of education successfully completed by the individuals of a given population. Furthermore, data on young people neither in employment nor in education and training -- NEET, early leavers from education and training and the labour status of young people by years since completion of highest level of education. Calculated as annual averages of quarterly EU Labour Force Survey data (EU-LFS).

## NUTS classification

@@

## Descriptive statistics

@@

## Part 1A: Sub-national GDP and GDP per Capita

### GDP per Capita Calculation

@@

### GDP per capita Portugal

@@

By looking at figure for Portugal, we can see that the GDP per capita in Portugal's regions appears to be fairly consistent. There is however some regional variability. We can see signs of urbanisazation where the regions around the big cities like Lisbon have a higher GDP per capita compared to some more rural areas. Since Lisbon is the capital of Portugal, there is probably a higher concentration of industries, making it a economic center (which again makes the GDP per capita higher).

@@

### GDP per capita France

@@

### GDP per capita Hungary

@@

### GDP per capita Slovakia

@@

### GDP per capita Denmark

@@

## Part 1B: Regional Inequity

### Gini Coefficient Calculation

In this part we will compute the population-weighted GDP Gini coefficient for each European NUTS2 region in our assigned countries.

@@

@@

## Part 2A: Growth and Inequity

@@

In this part of the paper, we will carry out a simple regression model and explore the effect of regional economic development, represented by 𝑦𝑗 (GDP per capita), on regional inequality, represented by 𝐺𝐼𝑁 𝐼𝑊𝑗. We will do this to gain an understanding of the connection between economic growth indicators such as GDP per capita and inequality might have. We will explore if higher GDP per capita may lead to less or greater inequality, and gain an understanding to what extent these variables are related.

Unlike the traditional Gini coefficient, which treats all individuals equally regardless of the population size of the region they reside in, the weighted Gini considers the population size of each region, assigning more weight to regions with larger populations.

In the context of regional inequality, this is particularly important because it ensures that the income disparities in more populous regions have a proportionally larger impact on the overall measure of inequality. For instance, if a country has one region with a very high level of income per capita but a small population, and another region with a lower level of income per capita but a large population, the weighted Gini coefficient would reflect the inequality experienced by a larger portion of the country's population, providing a more accurate picture of the national income distribution.

@@

$$
Y_i = β_1 + β_2X_i + ε_i
$$

$Y_i$ represent the dependent/ explained variable

$X_i$ represent the independent/ explanatory variable

$β_1$ represent the intercept/ constant

$β_2$ represent the slope coefficient

$ε_i$ represents the residuals or error in the prediction.

**Intercept (**β0​): represents the value of Y when X is 0.

@@

#### **Goodness of fit**

The goodness of fit in a simple linear regression model measures how well the regression line approximates the real data points. The regression line that best represents the data according to the least squares criterion, which minimizes the sum of the squared vertical distances of the points from the line.

R\^2 is a statistical measure that represents the proportion of the variance for the dependent variable that's explained by the independent variable. It ranges from 0 to 1. A higher R\^2 value indicates a better fit of the model to the data.

R\^2 = 0 The model does not explain any of the variability of the response data around its mean.

R\^2 = 1 The model explains all the variability of the response data around its mean.

If the assumptions for a simple linear regression are met, it indicates a good fit for the model. [Plotting the data can provide a visual indication of the goodness of fit.]{.underline} The points should fall around a straight line without clear patterns in the residuals. As it is crucial to verify these assumptions before proceeding with interpreting the results of the regression analysis.

**Residuals vs. Fitted Values Plot** to check for homoscedasticity and linearity. Ideally, this plot shows no pattern; the residuals are randomly scattered around the horizontal line at zero. If there's a pattern (like a curve or systematic spread of residuals), it suggests non-linearity or heteroscedasticity.

**Normal Q-Q Plot** to check if the residuals are approximately normally distributed. The points should fall roughly along a straight line. Deviations from a straight line suggest deviations from normality.

**Scale-Location Plot** is another way to check for homoscedasticity. The spread of the residuals should be roughly the same across all levels of the fitted values. If the spread increases or decreases with the fitted values, it suggests heteroscedasticity.

**Residuals vs. Leverage Plot** to identify influential observations that might unduly influence the regression. Points with high leverage or large residuals (outliers) can be identified. Cook's distance lines may be added to help identify points that have a large influence on the model.

## Part 2B: Exploring Other Determinants of Inequity

@@

Se data section for more specifications on the added population nut2, transport infrastructure and education.

@@

Multiple Linear Regression (MLR) extends simple linear regression to incorporate multiple explanatory variables, allowing us to examine how multiple factors impact a dependent variable. Choosing a data set from the year 2010 that consists of various regions (FRANCE?), with data on each region's economic indicators, demographic variables, and other factors. Our aim is to understand how these variables collectively affect regional inequality.

### **Assumptions**

**I. Linearity**

The relationship between the independent variable X and the dependent variable Y is linear. This means that changes in X are associated with proportional changes in Y. A straight line should provide a good fit to the data points when plotted on a graph.

You can create a scatter plot of Y versus X and visually inspect whether a straight line could well represent the relationship. Alternatively, you can plot the residuals versus the fitted values and check for any obvious patterns or non-linearity.

**II. Independence:**

Observations are independent of each other. The value of Y for one observation should not depend on the value of Y for any other observation.

This assumption is more about study design and data collection. Ensure that your observations are not correlated in time or space. For example, if you're analyzing economic data across regions, make sure that the regions are not influencing each other.

**III. Homoscedasticity (Equal Variance):**

The variance of the residuals (the errors) is constant across all levels of X. This means that the spread or "width" of the residuals should remain roughly the same across all values of the independent variable.

A plot of residuals versus fitted values should show a random scatter and not display any funnel-like shapes (wider at one end).

**IV. Normality of Residuals:**

The residuals (the differences between observed and predicted values) are normally distributed. This assumption is particularly important for hypothesis testing and creating confidence intervals.

A Quantile-Quantile (Q-Q) plot of the residuals can show if they follow a normal distribution. The points should fall roughly along a straight line.

**V. No Perfect Multicollinearity (Specific to Multiple Regression)**

In multiple regression settings, this assumes that no independent variable is a perfect linear function of any other independent variables. While this is more pertinent to multiple linear regression, it is crucial there because high correlation between independent variables can lead to unstable coefficient estimates.

Checking the variance inflation factor (VIF) for each variable; a VIF above 5-10 indicates a problematic amount of collinearity.

### Model specification

**Understanding the Coefficients**

Intercept beta_0 Represents the expected value of the dependent variable when all independent variables are set to zero. Interpretation is often nonsensical in multiple regression if there is no meaningful condition where all predictors are zero.

Slope Coefficients beta_1, beta_2, \..., \\beta_k \\))\*\*: Represent the expected change in the dependent variable for a one-unit change in the respective independent variable, holding all other variables constant.

### Ordinary Least Squares (OLS) Estimation

"The OLS method is employed to identify the best-fitting linear relationship between the dependent and independent variables, aiming to minimize the sum of squared residuals. This technique ensures that the estimations of the intercept (β0) and slope (β1) yield the least possible cumulative discrepancy between the actual and predicted values. The strength of OLS lies in its closed-form solution, providing a straightforward computation of coefficients directly from the data-set, without necessitating iterative procedures.

Furthermore, when the classical linear regression assumptions are met, OLS guarantees that the estimators are BLUE, ensuring their unbiasedness and efficiency. This is particularly crucial in econometric analysis, where the precision and reliability of parameter estimates are paramount for policy implications and economic interpretations. The foundation assumptions of linear regression, including linearity, independence, homoscedasticity, and normality of residuals, are prerequisites for OLS to attain these desirable properties. It is imperative, therefore, to conduct diagnostic tests and assess the validity of these assumptions to ensure the robustness of the OLS estimators used in our analysis."

#### **Model Validation with Cross-Sectional Data**

~~- \*\*Residual Analysis\*\*: Examine residuals for patterns to check for violations of linearity, independence, and homoscedasticity.~~

~~- \*\*Outlier and Influence Diagnostics\*\*: Identify regions with high leverage or large residuals to ensure they are not disproportionately affecting the model.~~

~~- \*\*Multicollinearity\*\*: Given that MLR includes multiple predictors, check for multicollinearity to ensure that the predictors are not too highly correlated with each other.~~

~~\#### \*\*Predictive Power and Goodness of Fit\*\*~~

~~- \*\*\\( R\^2 \\) and Adjusted \\( R\^2 \\)\*\*: Provide information on how well the independent variables explain the variation in the dependent variable.~~

~~- \*\*F-Test\*\*: Tests whether at least one of the independent variables is statistically significant in explaining variation in the dependent variable.~~

~~\#### \*\*Conclusion\*\*~~

~~- The multiple linear regression model, estimated using OLS, allows for the comprehensive analysis of how various factors impact regional inequality.~~

~~- Validating the assumptions of the OLS and properly interpreting the coefficients are vital steps in ensuring the reliability of the model.~~

~~- The use of cross-sectional data from the year 2010 ensures that the analysis is grounded in a specific time context, making the results more interpretable and applicable.~~

#### MLR Portugal

#### MLR Denmark

#### MLR France

#### MLR Hungary

#### MLR Slovakia
